---
title: "Honecker_BioData_V2_BassOysterFarmEnvironmental"
author: "Molly Honecker"
date: "2025-03-04"
output: html_document
---

# Load packages

```{r}

library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(Rmisc)
theme_set(theme_classic())

```

# Load data

```{r}
##### My tester 
R.version.string

BodyConditionTestCMAST <- read_excel("/Users/ameliamcdonald/Documents/Duke Grad/Bass/BioTeam/Body Condition 2025.xlsx", sheet = "CMAST")
BodyConditionTestWard <- read_excel("/Users/ameliamcdonald/Documents/Duke Grad/Bass/BioTeam/Body Condition 2025.xlsx", sheet = "WardCreek")
BodyConditionTestDUML <- read_excel("/Users/ameliamcdonald/Documents/Duke Grad/Bass/BioTeam/Body Condition 2025.xlsx", sheet = "DUML")

#not sure if this is needed 
#BodyCondition_Ratio_CMAST <- filter(BodyShellRatio_g, Site == "CMAST")
```
##### Previous Code 
#Mortality_All <- read_excel("/Users/ameliamcdonald/Documents/Duke Grad/Bass/BioTeam/Body Condition 2025.xlsx", sheet = "CMAST")
#Mortality_Continuous <- read_excel("Z:/Oyster Farm Environmental/OYSTERdataforR/BIOLOGICAL TEAM/Data/Processed/OysterMortalityData_Processed.xlsx", sheet = "OnlyContinuous")
#Mortality_Continuous_JulyDUMLMerged <- read_excel("Z:/Oyster Farm Environmental/OYSTERdataforR/BIOLOGICAL TEAM/Data/Processed/OysterMortalityData_Processed.xlsx", sheet = "OnlyContinuous_JulyDUMLMerged")
#Mortality_Continuous_PreventFranken <- read_excel("Z:/Oyster Farm Environmental/OYSTERdataforR/BIOLOGICAL TEAM/Data/Processed/OysterMortalityData_Processed.xlsx", sheet = "NoFrankenstein_OnlyContinuous")
#Mortality_FirstLast <-read_excel("Z:/Oyster Farm Environmental/OYSTERdataforR/BIOLOGICAL TEAM/Data/Processed/OysterMortalityData_Processed.xlsx", sheet = "FirstLast")


#Mortality_Continuous_NoMM <- filter(Mortality_Continuous, Strain != "MM")
#Mortality_Continuous_JulyDUMLMerged_NoMM <- filter(Mortality_Continuous_JulyDUMLMerged, Strain != "MM")
#Mortality_Continuous_DUML <- filter(Mortality_Continuous, Site == "DUML")
#Mortality_Continuous_DUML_NoMM <- filter(Mortality_Continuous_DUML, Strain != "MM")
#Mortality_Continuous_CMAST <- filter(Mortality_Continuous, Site == "CMAST")
#Mortality_Continuous_CMAST_NoMM <- filter(Mortality_Continuous_CMAST, Strain != "MM")
#Mortality_Continuous_PreventFranken_NoMM <- filter(Mortality_Continuous_PreventFranken, Strain != "MM")

Mortality_FirstLast_DUML <- filter(Mortality_FirstLast, Site == "DUML")
Mortality_FirstLast_CMAST <- filter(Mortality_FirstLast, Site == "CMAST")
#####
```

# SE/SD matrices

```{r}

## Calculate sd, se, and more for groups produced using "group" variables)

se <- summarySE(BodyConditionTestCMAST, 
                          measurevar="BodyShellRatio_g", 
                          groupvars=c("Date","Site"))

BodyConditionByStrain <- summarySE(BodyConditionTestCMAST, 
                          measurevar="BodyShellRatio_g", 
                          groupvars=c("Date","Color"))

```


```{r}

BodyConditionFunction <- function(dataframe) 
  ggplot(dataframe, aes(x = Date, y = BodyShellRatio_g,
                 color = Color)) +
  
  #Scatter plot (individual data points)
  geom_point(data = dataframe,
            aes(x = Date,
                y = BodyShellRatio_g,
                color = Color),
            size = 4,
            alpha = 0.075) +
  
  # Smoothed line (for trend)
  geom_line(data = dataframe,
            aes(x = Date,
                y = BodyShellRatio_g,
                color = Color),
            stat = "smooth",
            # method = "gam",
            size = 1.5,
            alpha = 0.75,
            se = F)


print(BodyConditionFunction)

```

```{r warning=FALSE}
DUML_plot <- BodyConditionFunction(BodyConditionTestDUML)
DUML_plot
```


```{r}
WARD_plot <- BodyConditionFunction(BodyConditionTestWard)
WARD_plot
```


```{r}
CMAST_plot <- BodyConditionFunction(BodyConditionTestCMAST)
CMAST_plot
```


# Mortality graph

```{r}
#Old code: plots with smoothed lines 

CMASTBodyCondition <- ggplot(BodyConditionTestCMAST, aes(x = Date, y = BodyShellRatio_g,
                 color = Color)) +
  
  #Scatter plot (individual data points)
  geom_point(data = BodyConditionTestCMAST,
            aes(x = Date,
                y = BodyShellRatio_g,
                color = Color),
            size = 4,
            alpha = 0.075) +
  
  # Smoothed line (for trend)
  geom_line(data = BodyConditionTestCMAST,
            aes(x = Date,
                y = BodyShellRatio_g,
                color = Color),
            stat = "smooth",
            # method = "gam",
            size = 1.5,
            alpha = 0.75,
            se = F) 

print(CMASTBodyShell)

```
```{r}
CMAST_plot <- BodyConditionFunction(BodyConditionTestCMAST)
CMAST_plot
```


# Mortality graph

```{r}
#Bar Chart: by farm, date as x axis plotting color

CMASTBodyCondition <- ggplot(BodyConditionTestCMAST, aes(x = as.factor(Date), y = BodyShellRatio_g, fill = Color)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "CMAST Body Condition by Genetic Strain",
    x = "Date",
    y = "Body Condition"
  ) +
  scale_fill_manual(
    values = c(
      "Purple" = "purple", 
      "Blue" = "blue",
      "Yellow" = "yellow",
      "Orange" = "orange"
    )
  )
  theme_minimal()


print(CMASTBodyCondition)
```
```{r}
#with function 
BodyConditionFunction3 <- function(dataframe, plot_title) {

BodyConditionFunction3 <- ggplot(dataframe, aes(x = as.factor(Date), y = BodyShellRatio_g, fill = Color)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = plot_title,
    x = "Date",
    y = "Body Condition"
  ) +
  scale_fill_manual(
    values = c(
      "Purple" = "purple", 
      "Blue" = "blue",
      "Yellow" = "yellow",
      "Orange" = "orange"
    )
  )
  theme_minimal()


print(BodyConditionFunction3)
}
```
```{r}
CMAST_plot3 <- BodyConditionFunction3(BodyConditionTestCMAST, "CMAST Body Condition by Date")
print(CMAST_plot3)
```
```{r}
DUML_plot3 <- BodyConditionFunction3(BodyConditionTestDUML, "DUML Body Condition by Date")
print(DUML_plot3)
```
```{r}
Ward_plot3 <- BodyConditionFunction3(BodyConditionTestWard, "Ward Body Condition by Date")
print(Ward_plot3)
```


```{r}
#Bar chart: by farm, color as x axis plotting month
BodyConditionTestCMAST$Month <- factor(BodyConditionTestCMAST$Month, levels = c("June", "July", "September"))

CMASTBodyCondition <- ggplot(BodyConditionTestCMAST, aes(x = Color, y = BodyShellRatio_g, fill = Month)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "CMAST Body Condition by Genetic Strain",
    x = "Color",
    y = "Body Condition"
  ) +
  scale_fill_manual(
    values = c(
     "June" = "lightblue", 
      "July" = "lightblue3",
      "September" = "lightblue4"
    )
  )
  theme_minimal()


print(CMASTBodyCondition)

```
```{r}
#Bar chart: by farm, color as x axis plotting month

BodyConditionFunction2 <- function(dataframe, plot_title) {
dataframe$Month <- factor(dataframe$Month, levels = c("June", "July", "September"))

BodyConditionFunction2 <- ggplot(dataframe, aes(x = Color, y = BodyShellRatio_g, fill = Month)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = plot_title,
    x = "Color",
    y = "Body Condition"
  ) +
  scale_fill_manual(
    values = c(
      "June" = "lightblue", 
      "July" = "lightblue3",
      "September" = "lightblue4"
    )
  )
  theme_minimal()


print(BodyConditionFunction2)
}
```
```{r}
CMAST_plot2 <- BodyConditionFunction2(BodyConditionTestCMAST, "CMAST Body Condition")
print(CMAST_plot2)
```
```{r}
DUML_plot2 <- BodyConditionFunction2(BodyConditionTestDUML, "DUML Body Condition")
print(DUML_plot2)
```
```{r}
Ward_plot2 <- BodyConditionFunction2(BodyConditionTestWard, "Ward Body Condition")
print(Ward_plot2)
```



  # Error bars (with SD)
  geom_pointrange(data = BodyConditionTestCMAST,
                aes(x = Date,
                    y = BodyShellRatio_g,
                    ymin = BodyShellRatio_g-sd,
                    ymax = BodyShellRatio_g+sd,
                    color = Site),
                linewidth = 1
                # alpha = 0.6
                ) +

  # Customize color scheme
  scale_color_manual(values = c("CMAST" = "#C84E00", "DUML" = "#0577B1")) +
  
  # Axis labels
  labs(x = "Date", 
       y = "Body Mass Ratio") +
  
  # Get rid of legend for color
  guides(color = "none") +
  
  # Facet
  
  # facet_wrap(vars(Site),
  #            nrow = 2) +
  
  # Adjust text sizes for axis and legend
  theme(
    axis.text = element_text(size = 14),   # Increase axis text size
    axis.title = element_text(size = 16),  # Increase axis title size
    legend.text = element_text(size = 14), # Increase legend text size
    legend.title = element_text(size = 16) # Increase legend title size
  )

# Uncomment the below to save the most recently produced graph to the specified path

# ggsave("PATH TO SAVE LOCATION",
#       width = 10,
#       height = 6)

```

